Concurrent Programming in Harmony

### 2020-12-01 book/paper.tex
Latest

\section*{Exercises}
\begin{problems}
\item See if you can implement barrier synchronization for \texttt{N} threads
with just three binary semaphores.  Busy waiting is not allowed.  How about just two?
(As always, the Little Book of Semaphores~\cite{Downey09} is a good resource
for solving synchronization problems with semaphores.
Look for the \emph{double turnstile}
\index{double turnstile}%
solution.)
\item Implement barrier synchronization with Mesa condition variables.
(You may want to use a double turnstile approach here as well.)
\end{problems}

>>>>

More generally, barrier synchronization can be abstracted as follows.
We want to create a \texttt{Barrier}$(n)$ object, with operations
\texttt{enter()} and \texttt{exit()}.  It is helpful to use
a roller coaster car with $n$ seats as a metaphor:
\begin{itemize}
\item the car cannot contain more than $n$ people;
\item the car won't take off until $n$ people are in the car;
\item no new people can enter the car until all $n$ people have left it.
\end{itemize}
Notice there are two different waiting conditions:
\begin{enumerate}
\item waiting for the car to empty out;
\item waiting for the car to fill up.
\end{enumerate}
Similarly, \texttt{enter}() blocks until $n$ threads have invoked it.
If more threads try to \texttt{enter}(), they will block until the
previous $n$ threads have invoked \texttt{exit}().

\autoref{fig:genbarrier} shows the code.
Method~\texttt{enter}() has two \texttt{wait} loops, one for each
waiting condition.
This is sometimes called \emph{double turnstile}.
Each loop uses a different condition variable.
The first loop waits while the barrier is fully occupied, while
the second waits for the barrier to fill up.
The $n^\mathit{th}$ thread to arrive at the barrier notifies the
$n-1$ threads that are waiting in the second loop.

Method \texttt{exit}() counts how many threads have left the
barrier.  When the last one leaves, it notifies all the threads
that are waiting in the first loop.

\section*{Exercises}
\begin{problems}
\item See if you can implement barrier synchronization for \texttt{N} threads
with just three binary semaphores.  Busy waiting is not allowed.  How about just two?
(As always, the Little Book of Semaphores~\cite{Downey09} is a good resource
for solving synchronization problems with semaphores.
Look for the \emph{double turnstile}
\index{double turnstile}%
solution.)
\end{problems}
